---
title: "BA 64060 - Assignment 1"
author: "Disney Maxwell"
date: "2025-09-06"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Dataset: Forbes – The Global 2000 Companies – 2025 

This dataset is from kaggle.com and contains financial data of the world's top 2000 companies.\
Dataset link: <https://www.kaggle.com/datasets/ellimaaac/forbes-the-global-2000-companies-2025>.

## 2. Import Dataset into R

Import financial data into R. 

```{r import}
library(readr)
forbes_2000_companies <- read_csv("Forbes_2000_Companies_2025.csv")
View(forbes_2000_companies)
```
## 3.A. Descriptive Statistics - Quantitative Variables 

i. Average Sales of 2000 Companies 

```{r avgSales}
sales_df <- forbes_2000_companies["Sales ($B)"]
mean_sales_value <- colMeans(sales_df)
print(paste0("The Mean sales value of the 2000 companies = ", mean_sales_value))
```

ii. Range Profit of 2000 Companies 

```{r rangeProfit}
profit_vector <- forbes_2000_companies[["Profit ($B)"]]
range_profit_value <- max(profit_vector, na.rm = TRUE) - min(profit_vector, na.rm = TRUE) 
print(paste0("The Range value for profit of the 2000 companies = ", range_profit_value))
```
## 3.B. Descriptive Statistics - Categorical Variables 

i. Number of Company Headquarters in different Countries
```{r headqrtrFreq}
headqrtr_freq <- table(forbes_2000_companies["Headquarters"])
print(paste0("Number of Headquarters across the different Countries:"))
print(headqrtr_freq)
```

ii. Country with the most number of Headquarters
```{r mostHeadqrtr}
most_freq_headqrtr <- names(which.max(headqrtr_freq))
print(paste0("Country with the most number of headquarters: ", most_freq_headqrtr))
```

## 4. Transformations - Convert Sales from Billion to Million dollar values
```{r millSalesvalues}
sales_millions_df <- sales_df * 1000
names(sales_millions_df) <- c("Sales ($M)")
first_10_companies_sales = head(sales_millions_df, 10)
print(paste0("First 10 Company Sales in million dollars"))
print(first_10_companies_sales)
```

## 5.A. Plot - Quantitative Variable - Top 5 Sales Values

```{r top5SalesCompanies}
sales_vector <- forbes_2000_companies[["Sales ($B)"]]
top_5_sales_vector <- sort(sales_vector, decreasing = TRUE)[1:5]
bar_x_labels <- c("1", "2", "3", "4", "5")

# Plot the vector as a bar chart
barplot(top_5_sales_vector, 
        names.arg = bar_x_labels, # Assign labels to bars
        main = "Top 5 Sales Values", # Set the main title of the plot
        xlab = "Indexes", # Set the x-axis label
        ylab = "Sales values", # Set the y-axis label
        col = "skyblue" # Set the color of the bars
        )
```

## 5.B. Scatter Plot - First 10 Companies - Sales and Profit 
```{r first10Companies}
first_10_companies <- forbes_2000_companies[1:10, c("Sales ($B)", "Profit ($B)")]
plot(first_10_companies$`Sales ($B)`, first_10_companies$`Profit ($B)`,
     xlab = "Sales ($B)", ylab = "Profit ($B)",
     main = "Scatter Plot of First 10 Companies - Sales and Profit")
```
